SMT-Boosted Security Types for Low-Level MPC

### 2025-11-07 toplas26/sty.tex
checkpoint

Both interactive and non-interactive judgements leverage the use
of SMT interpretations as described in Section \ref{section-smt}.
We introduce a \emph{share type} language $\sty$, classifying
\emph{share terms}. We represent share terms using standard notation
\cite{10.1007/978-3-030-68869-1_3}. A share term $[\phi_1,\phi_2]$
represents a secret-shared value where client 1 has $\phi_1$ and
client 2 has $\phi_2$, and this shared value can be reconstructed
as $\phi_1 + \phi_2$.\cnote{Comment on generation and
  reconstruction in general somewhere.} A share term $\pubx{\phi}$
denotes a publicly shared value.
$$
\begin{array}{rcl@{\hspace{10mm}}l}
  \sty &::=& \sharety \mid \pubty \mid \outty & \textit{share types} \\
  \stt &::=& [\phi,\phi] \mid \pubx{\phi} & \textit{share terms} 
\end{array}
$$ We proceed by formally describing the type analysis, then
discussing an illustrative example, and finally by demonstrating share
type soundness (Theorem \ref{theorem-sty}).

@@

The syntax-directed rules all known semi-homomorphic operations,
including addition of shares (\TirName{HEAdd}), and both addition and
multiplication of shares with a public value (\TirName{HEAddPub} and
\TirName{HEMultPub}, respectively) \cite{10.1007/978-3-030-68869-1_3}.
These rules also leverage SMT entailment in the \TirName{ShareEntails}
and \TirName{PubEntails} rules, since these operations should be
allowable up to algebraic equivalence in the context of a protocol.

@@

It is well-known that adding or subtracting a sample from a uniform
distribution in a finite field yields a value in a uniform
distribution, meaning that samples can be used as one-time-pads with
perfect secrecy \cite{barthe2019probabilistic,darais2019language}.

Of course, there are other methods for encrypting values in MPC
protocols-- but we can observe that many are algebraically equivalent
to this fundamental schema. For example, in Yao's Garbled Circuits
(YGC) in $\mathbb{F}_{2}$, the ``garbler'' represents encrypted wire
values as a random sample (for 1) or its negation (for 0), and shares
its secret input encoding with the ``evaluator'' by using
it as a selection bit. That is, assuming that client 2 is the garbler and
client 1 is the evaluator, we can define:
$$
\xassign{\mx{w}{1}}{\mux{\secret{w}}{\neg\flip{w}}{\flip{w}}}{2}
$$
where for all $\be_1,\be_2,\be_3$ (with $\neg$ denoting negation in $\mathbb{F}_2$):
$$
\mux{\be_1}{\be_2}{\be_3} \defeq (\neg\be_1 \ftimes \be_2) \fplus (\be_1 \ftimes \be_3)
$$
and letting this protocol be $\prog$ the following is valid:
\begin{mathpar}
%\toeq{\xassign{\mx{w}{1}}{\OT{\secret{w}}{1}{\neg\flip{w}}{\flip{w}}}{2}}
   \toeq{\prog} \models \mx{w}{1} \eop \neg\sx{w}{2} \fplus \rx{w}{2}
\end{mathpar}
This allows us to assign the same ciphertype to $\mx{w}{1}$ as above--
that is, $\cty{\rx{w}{2}}{\setit{\sx{w}{2}}}$.  Since this is a
one-time-pad scheme, it is also important that we ensure that samples
are used at most once for encryption. As in previous work 
\cite{darais2019language} we use type linearity to enforce this.

@@

>>>>

Both interactive and non-interactive judgements leverage the use
of SMT interpretations as described in Section \ref{section-smt}.
We introduce a \emph{share type} language $\sty$, classifying
\emph{share terms}. We represent share terms using standard notation
\cite{10.1007/978-3-030-68869-1_3}. A share term $[\phi_1,\phi_2]$
represents a secret-shared value where client 1 has $\phi_1$ and
client 2 has $\phi_2$, and this shared value can be reconstructed
as $\phi_1 + \phi_2$.\cnote{Comment on generation and
  reconstruction in general somewhere.} A share term $\pubx{\phi}$
denotes a publicly shared value.
$$
\begin{array}{rcl@{\hspace{10mm}}l}
  \sty &::=& \sharety \mid \pubty \mid \outty & \textit{share types} \\
  \stt &::=& [\phi,\phi] \mid \pubx{\phi} & \textit{share terms} 
\end{array}
$$
Intuitively, the meaning of share types $\sty$ are as follows.
When we ascribe the type $[\phi_1,\phi_2] : \sharety$, this
means that $[\phi_1,\phi_2]$ are shares of their
reconstruction, which we can represent as $\phi_1 + \phi_2$
as discussed in Section \ref{???}. 
In particular, each of $\phi_1$ and $\phi_2$ independently
reveal nothing about $\phi_1 + \phi_2$. Likewise,
$\pubx{\phi} : \outty$ means that $\phi$ has been
publicly \emph{revealed} via broadcast shares, and our
type system requires that $\phi \eop \phi_1 + \phi_2$ with
$[\phi_1,\phi_2] : \sharety$. This implies that the
process of revealing $\phi$ reveals nothing besides $\phi$.
The ascription $\pubx{\phi} : \pubty$ is similar, with the
additional meaning that $\phi$ is a publicly-known value
independent of any client inputs. This includes known
constant values used in protocols, and more interestingly
\emph{secure openings}. Secure openings are commonly used
to interactively compute public values that are masked with
OTPs secret-shared with clients in initialization protocols,
as for example in Beaver Triple multiplication which
we discuss in Section \ref{section-examples}.

We proceed by formally describing the type analysis, then
discussing an illustrative example, and finally by demonstrating share
type soundness in Theorem \ref{theorem-sty}. This result shows that
our analysis enforces noninterference modulo output, in contrast to
the previous system for $\minicat$ that only enforce a gradual release
property \cite{skalka-near-esop25}. By connecting this analysis with
protocol correctness properties enforced by the Hoare logic discussed
in Section \ref{section-smt}, the output can also be connected to an
intended ideal functionality.\cnote{Update intro grid accordingly.}

@@

The syntax-directed rules are all known semi-homomorphic operations,
including addition of shares (\TirName{HEAdd}), and both addition and
multiplication of shares with a public value (\TirName{HEAddPub} and
\TirName{HEMultPub}, respectively) \cite{10.1007/978-3-030-68869-1_3}.
These rules also leverage SMT entailment in the \TirName{ShareEntails}
and \TirName{PubEntails} rules, since these operations should be
allowable up to algebraic equivalence in the context of a protocol.

@@

As discussed in Section \ref{???}, a \emph{secret sharing scheme}
includes functions $\genshares$ and $\recon$ which generate
and reconstruct shares, respectively.
It is well-known that adding or subtracting a sample from a uniform
distribution in a finite field yields a value in a uniform
distribution, meaning that samples can be used as OTPs with
perfect secrecy \cite{barthe2019probabilistic,darais2019language}.
So we can define:
\begin{eqnarray*}
  \genshares(\phi) &\defeq& [\phi - \rx{y}{\cid_2}, \rx{y}{\cid_2}]\\
  \recon([\phi_1,\phi_2]) &\defeq& \phi_1 \fplus \phi_2
\end{eqnarray*}
where we require that $\genshares$ consumes the randomness
$\setit{\rx{y}{\cid_2}}$ for use as an OTP.

Of course, there are other methods for encrypting values in MPC
protocols that can be shown to be algebraically equivalent
to this fundamental schema. For example, in Yao's Garbled Circuits
(YGC) in $\mathbb{F}_{2}$, the ``garbler'' represents encrypted wire
values as a random sample (for 1) or its negation (for 0), and shares
its secret input encoding with the ``evaluator'' by using
it as a selection bit. That is, assuming that client 2 is the garbler and
client 1 is the evaluator, we can define:
$$
\xassign{\mx{w}{1}}{\mux{\secret{w}}{\neg\flip{w}}{\flip{w}}}{2}
$$
where for all $\be_1,\be_2,\be_3$ (with $\neg$ denoting negation in $\mathbb{F}_2$):
$$
\mux{\be_1}{\be_2}{\be_3} \defeq (\neg\be_1 \ftimes \be_2) \fplus (\be_1 \ftimes \be_3)
$$
and letting this protocol be $\prog$ the following is valid:
\begin{mathpar}
%\toeq{\xassign{\mx{w}{1}}{\OT{\secret{w}}{1}{\neg\flip{w}}{\flip{w}}}{2}}
   \toeq{\prog} \models \mx{w}{1} \eop \neg\sx{w}{2} \fplus \rx{w}{2}
\end{mathpar}
For this reason SMT entailment is also an important enabling technology
for the interactive rules.

@@

The interactive share typing rules are defined in Figure \ref{fig-styinteractive},
with judgements of the form $\eqs \vdash \Gamma_1,R_1, \stredx \Gamma_2,R_2$.
Each such judgement establishes a type classification for interactively-defined
variables-- that is, unicast and broadcast messages exchanged between
clients.

This allows us to assign the same ciphertype to $\mx{w}{1}$ as above--
that is, $\cty{\rx{w}{2}}{\setit{\sx{w}{2}}}$.  Since this is a
one-time-pad scheme, it is also important that we ensure that samples
are used at most once for encryption. As in previous work 
\cite{darais2019language} we use type linearity to enforce this.
